---
title: "BMU Fish Catch Data"
author: "Erika Gavenus"
date: "7/13/2018"
output:
  word_document: default
  pdf_document: default
  html_document: default
---

# Libraries
```{r libraries, message=FALSE, warning=FALSE}
  library(leaflet)
  library(dplyr)
  library(tidyr)
  library(ggplot2)
  library(DT)
  library(readxl)
  library(lubridate)
```

# Functions
``` {r functions}
drop.entry <- function(old.df, var, x) {
  
  new.df <- old.df[old.df[, var] != x, ]
  return(new.df)
}

keep.entry <- function(old.df, var, x) {
  
  new.df <- old.df[old.df[, var] == x, ]
  return(new.df)
}
```

# Setting working directory
```{r directory}
setwd("~/Box Sync/BMU Data")
```
# Importing data
```{r data import, message=FALSE, warning=FALSE}
Kitenyi.2014 <- read_excel("E-01_Kitenyi Beach_2014.xlsx")
Kitenyi.2015 <- read_excel("E-01_Kitenyi Beach_2015.xlsx")
Kitawi.2014 <- read_excel("E-02_Kitawi Beach_2014.xlsx")
Kitawi.2015 <- read_excel("E-02_Kitawi Beach_2015.xlsx")
Makira.2014 <- read_excel("E-03_Makira Beach_2014.xlsx")
Makira.2015 <- read_excel("E-03_Makira Beach_2015.xlsx")
Sena.2015 <- read_excel("E-04_Sena Beach_2015.xlsx")
Sena.2014 <- read_excel("E-04_Sena Beach_2014.xlsx")
Takawiri.2014<- read_excel("E-06_Takawiri Beach_2014.xlsx")
Takawiri.2015 <- read_excel("E-06_Takawiri Beach_2015.xlsx")
Mrongo.2014 <- read_excel("N-07_Mrongo Beach_2014.xlsx")
Mrongo.2015 <- read_excel("N-07_Mrongo Beach_2015.xlsx")
Likungu.2014 <- read_excel("N-08_Likungu Beach_2014.xlsx")
Likungu.2015 <- read_excel("N-08_Likungu Beach_2015.xlsx")
Yokia.2014<-read_excel("N-09_Yokia Beach_2014.xlsx")
Yokia.2014.omena <- read_excel("N-09_Yokia Beach _Omena_2014.xlsx")
Yokia.2015 <- read_excel("N-09_Yokia Beach_2015.xlsx")
Konyahero.2014 <- read_excel("N-10_Konyahero Beach_2014.xlsx")
Masisi.2014.omena <- read_excel("N-10_Masisi Beach _Omena_2014.xlsx")
Masisi.2014 <- read_excel("N-10_Masisi Beach_2014.xlsx")
Masisi.2015 <- read_excel("N-10_Masisi Beach_2015.xlsx")
Masisi.2014.ngege <- read_excel("N-10_Masisi Beach_Ngege_2014.xlsx")
Masisi.2014.ochangaa <- read_excel("N-10_Masisi Beach_Ochangaa_2014.xlsx")
Nyakweri.2014 <- read_excel("N-11_Nyakweri Beach_2014.xlsx")
Nyakweri.2015 <- read_excel("N-11_Nyakweri Beach_2015.xlsx")
Ugina.2014 <- read_excel("S-17_Ugina Beach_2014.xlsx")
Ugina.2015 <- read_excel("S-17_Ugina Beach_2015.xlsx")
Mauta.2014.omena <- read_excel("S-18_Mauta Beach _Omena_2014.xlsx")
Mauta.2014 <- read_excel("S-18_Mauta Beach_2014.xlsx")
Kasarani.2014 <- read_excel("S-19_Kasarani Beach_2014.xlsx")
Kasarani.2015 <- read_excel("S-19_Kasarani Beach_2015.xlsx")
Kiwari.2014 <- read_excel("W-12_Kiwari Beach_2014.xlsx")
Kiwari.2015 <- read_excel("W-12_Kiwari Beach_2015.xlsx")
Wakula.2014 <- read_excel("W-13_Wakula Beach_2014.xlsx")
Nyawalongo.2014 <- read_excel("W-14_Nyawalongo Beach_2014.xlsx")
Nyawalongo.2015 <- read_excel("W-14_Nyawalongo Beach_2015.xlsx")
Nyawalongo.2014.omena <- read_excel("W-14_Nyawalongo Beach_Omena_2014.xlsx")
Mulundu.2014 <- read_excel("W-15_Mulundu Beach_2014.xlsx")
Mulundu.2014.omena <- read_excel("W-15_Mulundu Beach_Omena_2014.xlsx")
Tiko.2014 <- read_excel("W-16_Tiko Beach_2014.xlsx")
```

# Cleaning column names
``` {r column names}
Sena.2014.clipped<-subset(Sena.2014, select = -c(Small_Wt, Med_Wt, Large_Wt))
Sena.2014<-Sena.2014.clipped
Wakula.2014$Total<-NA
Wakula.2014$Note<-NA

colnames(Masisi.2014.omena)[colnames(Masisi.2014.omena) == 'Troughs'] <- 'Weight'
colnames(Mauta.2014.omena)[colnames(Mauta.2014.omena) == 'Troughs'] <- 'Weight'
colnames(Mulundu.2014.omena)[colnames(Mulundu.2014.omena) == 'Troughs'] <- 'Weight'
colnames(Nyawalongo.2014.omena)[colnames(Nyawalongo.2014.omena) == 'Troughs'] <- 'Weight'
colnames(Yokia.2014.omena)[colnames(Yokia.2014.omena) == 'Troughs'] <- 'Weight'

```

# Adding columns for BMU, region, fish species
```{r adding columns}
Kasarani.2014$bmu<-"Kasarani"
Kasarani.2014$fish<-"mbuta"
Kasarani.2014$region<-"S"

Kasarani.2015$bmu<-"Kasarani"
Kasarani.2015$fish<-"mbuta"
Kasarani.2015$region<-"S"

Kitawi.2014$bmu<-"Kitawi"
Kitawi.2014$fish<-"mbuta"
Kitawi.2014$region<-"E"

Kitawi.2015$bmu<-"Kitawi"
Kitawi.2015$fish<-"mbuta"
Kitawi.2015$region<-"E"

Kitenyi.2014$bmu<-"Kitenyi"
Kitenyi.2014$fish<-"mbuta"
Kitenyi.2014$region<-"E"

Kitenyi.2015$bmu<-"Kitenyi"
Kitenyi.2015$fish<-"mbuta"
Kitenyi.2015$region<-"E"

Kiwari.2014$bmu<-"Kiwari"
Kiwari.2014$fish<-"mbuta"
Kiwari.2014$region<-"W"

Kiwari.2015$bmu<-"Kiwari"
Kiwari.2015$fish<-"mbuta"
Kiwari.2015$region<-"W"

Konyahero.2014$bmu<-"Konyahero"
Konyahero.2014$fish<-"mbuta"
Konyahero.2014$region<-"N"

Likungu.2014$bmu<-"Likungu"
Likungu.2014$fish<-"mbuta"
Likungu.2014$region<-"N"

Likungu.2015$bmu<-"Likungu"
Likungu.2015$fish<-"mbuta"
Likungu.2015$region<-"N"

Makira.2014$bmu<-"Makira"
Makira.2014$fish<-"mbuta"
Makira.2014$region<-"E"

Makira.2015$bmu<-"Makira"
Makira.2015$fish<-"mbuta"
Makira.2015$region<-"E"

Masisi.2014$bmu<-"Masisi"
Masisi.2014$fish<-"mbuta"
Masisi.2014$region<-"N"


Masisi.2014.ochangaa$bmu<-"Masisi"
Masisi.2014.ochangaa$fish<-"ochangaa"
Masisi.2014.ochangaa$region<-"N"

Masisi.2014.ngege$bmu<-"Masisi"
Masisi.2014.ngege$fish<-"ngege"
Masisi.2014.ngege$region<-"N"

Masisi.2014.omena$bmu<-"Masisi"
Masisi.2014.omena$fish<-"omena"
Masisi.2014.omena$region<-"N"

Masisi.2015$bmu<-"Masisi"
Masisi.2015$fish<-"mbuta"
Masisi.2015$region<-"N"

Mauta.2014$bmu<-"Mauta"
Mauta.2014$fish<-"mbuta"
Mauta.2014$region<-"S"

Mauta.2014.omena$bmu<-"Mauta"
Mauta.2014.omena$fish<-"omena"
Mauta.2014.omena$region<-"S"

Mrongo.2014$bmu<-"Mrongo"
Mrongo.2014$fish<-"mbuta"
Mrongo.2014$region<-"N"

Mrongo.2015$bmu<-"Mrongo"
Mrongo.2015$fish<-"mbuta"
Mrongo.2015$region<-"N"

Mulundu.2014$bmu<-"Mulundu"
Mulundu.2014$fish<-"mbuta"
Mulundu.2014$region<-"W"

Mulundu.2014.omena$bmu<-"Mulundu"
Mulundu.2014.omena$fish<-"omena"
Mulundu.2014.omena$region<-"W"

Nyakweri.2014$bmu<-"Nyakweri"
Nyakweri.2014$fish<-"mbuta"
Nyakweri.2014$region<-"N"

Nyakweri.2015$bmu<-"Nyakweri"
Nyakweri.2015$fish<-"mbuta"
Nyakweri.2015$region<-"N"

Nyawalongo.2014$bmu<-"Nyawalongo"
Nyawalongo.2014$fish<-"mbuta"
Nyawalongo.2014$region<-"W"

Nyawalongo.2014.omena$bmu<-"Nyawalongo"
Nyawalongo.2014.omena$fish<-"omena"
Nyawalongo.2014.omena$region<-"W"

Nyawalongo.2015$bmu<-"Nyawalongo"
Nyawalongo.2015$fish<-"mbuta"
Nyawalongo.2015$region<-"W"

Sena.2014$bmu<-"Sena"
Sena.2014$fish<-"mbuta"
Sena.2014$region<-"E"

Sena.2015$bmu<-"Sena"
Sena.2015$fish<-"mbuta"
Sena.2015$region<-"E"

Takawiri.2014$bmu<-"Takawiri"
Takawiri.2014$fish<-"mbuta"
Takawiri.2014$region<-"E"

Takawiri.2015$bmu<-"Takawiri"
Takawiri.2015$fish<-"mbuta"
Takawiri.2015$region<-"E"

Tiko.2014$bmu<-"Tiko"
Tiko.2014$fish<-"mbuta"
Tiko.2014$region<-"W"

Ugina.2014$bmu<-"Ugina"
Ugina.2014$fish<-"mbuta"
Ugina.2014$region<-"S"

Ugina.2015$bmu<-"Ugina"
Ugina.2015$fish<-"mbuta"
Ugina.2015$region<-"S"

Wakula.2014$bmu<-"Wakula"
Wakula.2014$fish<-"mbuta"
Wakula.2014$region<-"W"

Yokia.2014$bmu<-"Yokia"
Yokia.2014$fish<-"mbuta"
Yokia.2014$region<-"N"

Yokia.2014.omena$bmu<-"Yokia"
Yokia.2014.omena$fish<-"omena"
Yokia.2014.omena$region<-"N"

Yokia.2015$bmu<-"Yokia"
Yokia.2015$fish<-"mbuta"
Yokia.2015$region<-"N"

```

# Compiling data into single dataframe
``` {r binding}
catchdata.full<- rbind(Kasarani.2014, Kasarani.2015, Kitawi.2014, Kitawi.2015, Kitenyi.2014, Kitenyi.2015, Kiwari.2014, Kiwari.2015, Konyahero.2014, Likungu.2014, Likungu.2015, Makira.2014, Makira.2015, Masisi.2014, Masisi.2014.ngege, Masisi.2014.ochangaa, Masisi.2014.omena, Masisi.2015, Mauta.2014, Mauta.2014.omena, Mrongo.2014, Mrongo.2015, Mulundu.2014, Mulundu.2014.omena, Nyakweri.2014, Nyakweri.2015, Nyawalongo.2014, Nyawalongo.2014.omena, Nyawalongo.2015, Sena.2014, Sena.2015, Takawiri.2014, Takawiri.2015, Tiko.2014, Ugina.2014, Ugina.2015, Wakula.2014, Yokia.2014, Yokia.2014.omena, Yokia.2015)
```

# Limiting to mbuta data
``` {r mbuta.data}
mbutacatch<-keep.entry(catchdata.full, 'fish', 'mbuta')
```

# Cleaning date errors
``` {r date errors}
mbutacatch$Date<-as_date(mbutacatch$Date)
mbutacatch$Date[which (mbutacatch$Date=="2104-06-02")]<- "2014-06-02"
mbutacatch$Date[which (mbutacatch$Date=="2104-01-12")]<- "2014-01-12"
mbutacatch$Date[which (mbutacatch$Date=="2103-10-22")]<- "2013-10-22" 

```

# Summarizing mbuta catch and rates by date and beach
``` {r totalcatch.date.beach}
beach.totals <- mbutacatch %>% 
    group_by(Date, bmu) %>% 
    summarise(daily.total=sum(Weight), rate.avg=mean(Rate))
```

# Graphing daily mbuta catch by bmu
``` {r bmu daily catch graph}
daily.catch.bmu = ggplot(data=beach.totals, aes(x=Date, y=daily.total, color=bmu, fill=bmu)) +  
    geom_line () +
    theme_bw() + theme() + labs(x="Date", y="Total Catch (kg)") + ggtitle("Daily Catch of Mbuta (kg) by BMU")
print(daily.catch.bmu)
```

# Summarizing mbuta catch and rates by date and region
``` {r totalcatch.date.region}
region.totals <- mbutacatch %>% 
    group_by(Date, region) %>% 
    summarise(daily.total=sum(Weight), rate.avg=mean(Rate))
```

# Graphing daily mbuta catch by region
``` {r regional daily catch graph}
daily.catch.region = ggplot(data=region.totals, aes(x=Date, y=daily.total, color=region, fill=region)) +  
    geom_line () +
    theme_bw() + theme() + labs(x="Date", y="Total Catch (kg)") + ggtitle("Daily Catch of Mbuta (kg) by Region")
print(daily.catch.region)
```

# Summarizing mbuta catch and rates by date
``` {r totalcatch.date}
totalcatch <- mbutacatch %>% 
    group_by(Date) %>% 
    summarise(daily.total=sum(Weight), rate.avg=mean(Rate))
```

# Graphing daily mbuta catch
``` {r daily catch graph}
daily.catch = ggplot(data=totalcatch, aes(x=Date, y=daily.total)) +  
    geom_line () +
    theme_bw() + theme() + labs(x="Date", y="Total Catch (kg)") + ggtitle("Daily Catch of Mbuta (kg)")
print(daily.catch)
```

# Daily catch by number of beaches reporting
